Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
edits up to effective hypersurfaces

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-span}
\sum_{i=1}^{n} c_i \cdot \bida_i \cdot f_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

We first restrict attention to the case $n=1$, so $D = \alpha D_i$.

\ssec{Effective Divisors of Hypersurfaces on Projective Space}
\label{ssec:proj-one-point}

>>>>

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-span}
\sum_{i=1}^{n} c_i \cdot \bida_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

\ssec{Effective Divisors on Projective Space}
\label{ssec:proj-one-point}

In this subsection, we restrict attention to the case of effective divisors.
